---
layout: post
title: Flake it till you make it
subtitle: Excerpt from Soulshaping by Jeff Brown
bigimg: /img/path.jpg
output: 
  html_document:
  md_document:
    variant: markdown
---

The [twitteR](https://github.com/geoffjentry/twitteR) allows easy access to the Twitter API within R. It allows users
to extract subsets of Twitter data for their own purposes. The purpose of the article
is to give tips on setting up twitteR.


Twitter (and TwitteR) use OAuth for security. What this means is that our username and password
aren't transmitted. To implement OAuth, you first create an [App](https://apps.twitter.com/). After logging
in, click 

> Create New App

Fill in the Application details:

  * __Name__: A new unique identifer
  * __Description__: My first twitter app
  * __Website__: http://www.example.com
  * __Callback URL__: http://127.0.0.1:1410

After reading the Developer argeement, click ___Yes__ and Create your Twitter application.

Next, click on `Keys and Access Tokens` and change the Access level from `Read and write`
to `Read, write and access direct messages`. Next copy the `Consumer Key and Secret` into R:

```{r}
api_key = "aLongString"
api_secret = "AnotherLongString"
```

Then run
```{r eval=FALSE}

library("twitteR")
setup_twitter_oauth(api_key, api_secret)
```
This will open a page on your browser. You should see output similar to
```{r eval=FALSE}
#[1] "Using browser based authentication"
#Use a local file ('.httr-oauth'), to cache OAuth access credentials between R sessions?

#1: Yes
#2: No

#Selection: Yes
#Adding .httr-oauth to .gitignore
#Waiting for authentication in browser...
#Press Esc/Ctrl + C to abort
#Authentication complete.
```

```{r}
tweets = searchTwitter("#rstats", n = 5, lang = "en")
strip_retweets(tweets, strip_manual=TRUE, strip_mt=TRUE)[[1]]
```

http://thinktostart.com/twitter-authentification-with-r/
http://davetang.org/muse/2013/04/06/using-the-r_twitter-package/

